---
title: "File Input and Output"
author: "Garrett B. Lucas"
date: "3/1/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

How do we even get data into R in the first place?

file.path() or here::here() for file pathway

```{r}
write_csv(gap_asia_2007, 
          here::here("data", "gap_asia_2007.csv")
)
```

Reading Data from Disk:

```{r}
dat <- read_csv(here::here("data", "gap_asia_2007.csv"))
dat
```

cols function controls the types of columns. (Below is the same as above, but the columns are specified). Its always a good idea to be explicit and explain what type of variable each column is.
Dbl- Numerical
Chr- Non-Numerical

```{r}
dat <- read_csv(
  here::here("data", "gap_asia_2007.csv"),
  col_types = cols(
    country = col_factor(),
    continent = col_factor(),
    year = col_date(format = "%Y"),
    .default = col_double() # all other columns as numeric (double)
  )
)
dat
```

Also include "na", or which values in the data should be considered na. Below considers blank cells, no response, and NA as na variables.

```{r}
dat <- read_csv(
  here::here("data", "gap_asia_2007.csv"),
  col_types = cols(
    country = col_factor(),
    continent = col_factor(),
    year = col_date(format = "%Y"),
    .default = col_double() # all other columns as numeric (double)
  ),
  na = c("", "NA", -99, "No response")
)
dat
```

1) File Path
2) Types of Data
3) NAs

Excel being imported is simple with the readxl command. (dont trust excel with your data, use google sheets.)

```{r}

```
